%% 根据自己实际情况增加或删减
\begin{publications}
\noindent
\textbf{发表与接收论文}
\renewcommand{\labelenumi}{[\arabic{enumi}]}
\begin{enumerate}
\item \textbf{Bin Liu}, Bang Wang*. Bayesian Negative Sampling for Recommendation. \textit{IEEE International Conference on Data Engineering} (ICDE), 2023, 749-761. (\textbf{CCF-A}; 署名单位：华中科技大学)

\item \textbf{Bin Liu}, Bang Wang*. Pairwise Learning for Personalized Ranking with Noisy Comparisons. \textit{Information Sciences}, 2022, 623: 242-257. (\textbf{中科院一区TOP}；IF=8.1; 署名单位：华中科技大学)
\item \textbf{Bin Liu}, Tao Chen, Peng Jia, Lin Wang*. Effective public service delivery supported by time-decayed Bayesian personalized ranking. \textit{Knowledge-Based Systems}, 2020, 206: 106376. (\textbf{中科院一区TOP}；IF=8.8; 署名单位：华中科技大学)
\end{enumerate}

\noindent
\textbf{正在投稿论文}
\begin{enumerate}
\item \textbf{Bin Liu}, Bang Wang*. Bayesian self-supervised Contrastive Learning. \textit{arXiv preprint arXiv: 2301.11673}, 2023. (\textbf{AAAI} under review (Phase 2); 署名单位：华中科技大学)
\item \textbf{Bin Liu}, Qin Luo, Bang Wang*. Debiased Pairwise Learning from Positive-Unlabeled Implicit Feedback. \textit{arXiv preprint arXiv:2307.15973}, 2023. (\textbf{IEEE Transactions on Knowledge and Data Engineering} under review; 署名单位：华中科技大学)
\item \textbf{Bin Liu}, Eerjia Chen, Bang Wang*. Reducing Popularity Bias in Recommender Systems through AUC-Optimal Negative Sampling. \textit{arXiv preprint arXiv:2306.01348}, 2023. (\textbf{AAAI} under review (Phase 2); 署名单位：华中科技大学)

\end{enumerate}
%\textbf{科技奖励}
%\renewcommand{\labelenumi}{[\arabic{enumi}]}
%\begin{enumerate}
%\item 全部作者的姓名全称，本人的名字加粗. 题目. 国家级/省部级科技类奖，获奖年
%\item 全部作者的姓名全称，本人的名字加粗. 题目. 国际/国内竞赛类奖，获奖年
%\end{enumerate}
\end{publications}
